### 2022-01-06 src-latex/main-paper.tex
Suggestion of structure
Update main-paper.tex
Proposition d'intro: à corriger/compléter/amender...

% \usetikzlibrary{shadings,decorations,decorations.pathreplacing}
\renewcommand\UrlFont{\color{blue}\rmfamily}

@@

\begin{document}
%
\title{A Benchmark of Named Entity Recognition Approaches in Historical Documents\\
Application to 19$^{th}$ Century French Directories%
% FIXME "19$^{th}$-Century" with dash?
\thanks{This work was supported by the French National Research Agency:
Project SoDuCo, grant ANR-18-CE38-0013. ???FIXME need acknowledgments for directory sources???}}
%
\titlerunning{A Benchmark of NER Approaches in Historical Documents}
%
\author{LISTE PROVISOIRE PAR ORDRE ALPHA\\
%
Nathalie Abadie\inst{1}\orcidID{0000-1111-2222-3333} \and
Edwin Carlinet\inst{2,3}\orcidID{1111-2222-3333-4444} \and
Joseph Chazalon\inst{3}\orcidID{2222--3333-4444-5555} \and
Bertrand Duménieu\inst{3}\orcidID{2222--3333-4444-5555}}
%
\authorrunning{F. Author et al.}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{INST, CITY, COUNTRY \and
INST, CITY, COUNTRY\\
\email{example@example.com}\\
\url{https://www.example.com} \and
INST, CITY, COUNTRY\\
\email{example@example.com}}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
\input{parts/00-abstract.tex}
\keywords{Historical documents \and Natural Language Processing \and Named Entity Recognition \and OCR noise \and Annotation cost.}
\end{abstract}

@@

>>>>

% \usetikzlibrary{shadings,decorations,decorations.pathreplacing}
\renewcommand\UrlFont{\color{blue}\rmfamily}
% FIXME proper hyperref setup

@@

\begin{document}
%
\title{A Benchmark of Named Entity Recognition Approaches in Historical Documents\\
Application to 19$^{th}$ Century French Directories%
% FIXME "19$^{th}$-Century" with dash?
\thanks{This work was supported by the French National Research Agency:
Project SoDuCo, grant ANR-18-CE38-0013. ???FIXME need acknowledgments for directory sources???}}
%
\titlerunning{A Benchmark of NER Approaches in Historical Documents}
%
\author{LISTE PROVISOIRE PAR ORDRE ALPHA\\
%
Nathalie Abadie\inst{1}\orcidID{0000-0001-8741-2398} \and
Edwin Carlinet\inst{2,3}\orcidID{1111-2222-3333-4444} \and
Joseph Chazalon\inst{3}\orcidID{2222--3333-4444-5555} \and
Bertrand Duménieu\inst{3}\orcidID{2222--3333-4444-5555}}
%
\authorrunning{F. Author et al.}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{INST, CITY, COUNTRY \and
INST, CITY, COUNTRY\\
\email{example@example.com}\\
\url{https://www.example.com} \and
INST, CITY, COUNTRY\\
\email{example@example.com}}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
\input{parts/00-abstract.tex}
\keywords{Historical documents \and Natural Language Processing \and Named Entity Recognition \and OCR noise \and Annotation cost.}
\end{abstract}

@@

\section*{Thanks}
Laureencia for annotations, Guillaume Thomas for data processing?

### 2022-01-06 src-latex/parts/10-introduction.tex
Proposition d'intro: à corriger/compléter/amender...

% General context + scope limitation
NER is a fundamental step of the information extraction pipeline when processing historical listing, like directories for instance.
OCRed text is not sufficient to build a high level, semantic view of the collection under study;
indeed, being able to properly tag text tokens unlocks the ability to relate entities,
and provide colleagues from digital humanities with useful databases.

% Problematic
NER requires training data, and we have little training data for historical documents.
OCR is also noisy.
How can we extract reliable elements in this context?

% Contributions
\begin{enumerate}
    \item review state of the art regarding NER for historical documents (section 2)
    \item dataset (section 3)
    \item list usable techniques (more to evangelize the DAR community) (section 4)
    \item compare their performance under various training set sizes (thus annotation costs) and OCR noise levels (thus OCR choices/luck/availability) (sec. 5 and 6)
    \item suggest smart way to leverage side knowledge like lexicons (FIXME iif this works well, maybe a conclusion instead)
\end{enumerate}

>>>>

% General context + scope limitation

Heritage institutions such as archives or libraries preserve large collections of textual documents, increasingly digitized or natively digital. They disseminate these digital documents on the Web to enable users to access them, thus making these knowledge sources available for numerous applications in the fields of social sciences, environmental studies, education, etc. Fostering information search and reuse within these collections requires two steps. The first step, called Optical Character Recognition (OCR), aims at transforming the images produced during the digitization of the documents into usable digital texts. However, OCRed text is not sufficient to build a high level, semantic view of the collection under study. A second step thus aims at recognizing in these texts the pieces of information most likely to be searched for by users, such as named entities (people, organisations, dates, places, etc.). Indeed, being able to properly tag text tokens unlocks the ability to relate entities, and provide colleagues from digital humanities with useful databases.

In the case of historical textual documents, OCR and Named Entity Recognition (NER) tasks are particularly challenging. Indeed, even in the case of printed historical documents, the OCR approaches used for modern documents fail due to many reasons: poor scan quality, irregular printing baseline, printing artifacts, old fonts, complex page layouts, special glyphs, etc. \textcolor{green}{Phrases précedentes à reprendre avec les termes adéquats + ajouter une ref.} On the other hand, NER approaches developed for modern texts suffer from the time gap between the contemporary entities and the entities mentioned in older texts and from the errors introduced into the text by the OCR step.\textcolor{green}{ajouter une ref}

In this article we focus on a particular type of historical textual sources, as they provide very rich, semi-structured and time-stamped information about people, their activities and their locations: the directories. In a recent work to locate potentially polluted urban soils \cite{bell2020automated}, historical directories are leveraged to identify and locate all the gas stations that were established in the city of Providence during the 20th century. The proposed information extraction pipeline includes OCR, entries identification and business type and address extraction. In the same way, we aim at producing structured spatio-temporal data from the information contained in the 19th century Parisian trade directories. These directories are produced by various publishers, organised according to different indexing methods (by name, by activity, by location, etc.), printed in different layouts, with different fonts, etc. We therefore investigate various state-of-the-art OCR and NER approaches independently and in combination, to assess their usability on our historical directories corpus. Thus, the contributions of this article are the following:
% Contributions
\begin{enumerate}
    \item review state of the art regarding NER for historical documents (section 2)
    \item dataset (section 3)
    \item list usable techniques (more to evangelize the DAR community) (section 4)
    \item compare their performance under various training set sizes (thus annotation costs) and OCR noise levels (thus OCR choices/luck/availability) (sec. 5 and 6)
    \item suggest smart way to leverage side knowledge like lexicons (FIXME iif this works well, maybe a conclusion instead)
\end{enumerate}

### 2022-01-06 src-latex/parts/30-dataset.tex
Proposition d'intro: à corriger/compléter/amender...

TODO interesting points to detail:
\begin{itemize}
    \item origin
    \item illustration (maybe in first figure)
    \item various numbers (of original data and of the sample we extracted)
    \item statistics about entry lengths (histogram of number of char / tokens par entry)
    \item entities we considered and how we annotated them (how we assessed gold's quality)
    \item OCR data (automated systems and clean text + how we created the "clean" text + whether/how we assessed clean text's quality)
    \item \dots
\end{itemize}

>>>>

TODO interesting points to detail:
\begin{itemize}
    \item origin
    \item illustration (maybe in first figure)
    \item various numbers (of original data and of the sample we extracted) : number of entries per pages, variety of directory entries, \dots
    \item statistics about entry lengths (histogram of number of char / tokens par entry)
    \item entities we considered and how we annotated them (how we assessed gold's quality)
    \item statistics on the distribution of entities across directories and their arrangement within entries
    \item OCR data (automated systems and clean text + how we created the "clean" text + whether/how we assessed clean text's quality)

    \item \dots
\end{itemize}

### 2022-01-06 src-latex/parts/40-ner-approaches.tex
Proposition d'intro: à corriger/compléter/amender...

% !TeX root = ../main-paper.tex
\section{Dataset}

>>>>

% !TeX root = ../main-paper.tex
\section{NER Approaches}

\begin{itemize}
    \item regexp
    \item Spacy CNN with tok2vec
    \item CamemBERT finetuned with wikiner-fr + gold dataset (clean texts)
    \item CamemBERT finetuned with wikiner-fr + gold dataset (clean texts + pretrained on raw texts (Pero ? Kraken ? Both ?)
\end{itemize}

% Other possibilities:
1. Traditional ML based:
    Conditional Random Fields (CRF) - https://pypi.org/project/sklearn-crfsuite/
    Maximum-entropy Markov model

2. Neural Networks based:
    LSTMs, bi-LSTM - https://github.com/flairNLP/flair
    CNNs (SpaCy uses CNN based architecture)
    Transformers (Spacy has recently launched it) - https://spacy.io/universe/project/spacy-transformers

### 2022-01-06 src-latex/parts/50-xp-setup.tex
Proposition d'intro: à corriger/compléter/amender...

\subsection{Metrics}
\begin{itemize}
    \item NER quality: accuracy? precision? recall? f-score? (idea: stick with standard spacy/sklearn metrics) \textcolor{blue}{Dans la plupart des articles que j'ai lus, c'est toujours une F-mesure qui est utilisée et parfois la précision et le rappel sont ajoutés.}
    \item OCR accuracy: outputs from standard tools (beware of unknown glyphs: do we need any special care?) \textcolor{blue}{Pour l'instant, j'ai croisé: CER, WER, Levenshtein et Jaccard. J'ai encore un état de l'art à lire sur le sujet: il y aura probablement d'autres mesures mentionnées.}

@@

\subsection{Training procedure}
train/val/test splits, number of runs, etc.

>>>>

\subsection{Metrics}
\begin{itemize}
    \item NER quality: Precision, Recall, F1 score (idea: stick with standard spacy/sklearn metrics) \textcolor{blue}{Dans la plupart des articles que j'ai lus, c'est toujours une F-mesure qui est utilisée et parfois la précision et le rappel sont ajoutés.}
    \item OCR accuracy: outputs from standard tools (beware of unknown glyphs: do we need any special care?) \textcolor{blue}{Pour l'instant, j'ai croisé: CER, WER, Levenshtein et Jaccard. J'ai encore un état de l'art à lire sur le sujet: il y aura probablement d'autres mesures mentionnées.}

@@

\subsection{Training procedure}
5-fold CV for all models.
\begin{itemize}
\item Spacy CNN : no limits on the number of epochs, instead use patience=1600.
\item CamemBERT : 3 epochs.
\end{itemize}
